---
title: "Compleat Database e-letter in Science Signaling"
author: "Savas Paragamian"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  bookdown::pdf_book: default
  html_document: default
  bookdown::gitbook: default
  word_document: default
documentclass: article
link-citations: yes
bibliography: Networks Centralities and Negative Weigths.bib
site: bookdown::bookdown_site
biblio-style: apalike
---

# Compleat Database


The data

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(scales)
library(knitr)
library(readr)
```

From the COMPLEAT database and online tool we downloaded all the protein complexes of drosophila. The complexes are both from literature and predicted with a variety of tools. The methology is explained in the paper as well as in the [online portal](http://www.flyrnai.org/compleat/) [@Vinayagam2013a]. 

```{r,warning=FALSE,message=FALSE}
# drosophila 

compleat_drosophila <- read.delim(file = "Data/drosophila_complexes_compleat0.txt",header = F,sep = "\t")
compleat_drosophila <- compleat_drosophila[!(is.na(compleat_drosophila$V2)),] # remove empty rows
# homo sapiens
compleat_homo <- read.delim(file = "Data/compleat1_homo.txt",header = F,sep = "\t")
compleat_homo <- compleat_homo[!(is.na(compleat_homo$V2)),] # remove empty rows
# yeast
compleat_yeast <- read.delim(file = "Data/compleat2_yeast.txt",header = F,sep = "\t")
compleat_yeast <- compleat_yeast[!(is.na(compleat_yeast$V2)),] # remove empty rows

```



# Distribution of the size of complexes

Plot the distributions

```{r}
# drosophila
compleat_drosophila_complex_distribution <- compleat_drosophila %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_drosophila_complex_distribution)[1] <- "complex_with_k_proteins"

# homo sapiens
compleat_homo_complex_distribution <- compleat_homo %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_homo_complex_distribution)[1] <- "complex_with_k_proteins"

# yeast
compleat_yeast_complex_distribution <- compleat_yeast %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_yeast_complex_distribution)[1] <- "complex_with_k_proteins"


```

Whole distributions. 

```{r}
ggplot()+
  geom_point(data = compleat_drosophila_complex_distribution, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="Fruit fly"))+
  geom_point(data = compleat_homo_complex_distribution, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="Human"))+
  geom_point(data = compleat_yeast_complex_distribution, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="Yeast"))+
  ggtitle("Distribution of Proteins in Complexes from COMPLEAT")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Fruit fly"="orange1", "Human"= "gray48", "Yeast"= "mediumaquamarine"), name="Organisms")+
  labs(x="number of proteins", y="number of complexes")+
  theme_bw()
```

The human dataset doesn't have information on the methods that the complexes were discovered.

Literature VS Predicted

```{r}
## Methods
# drosophila
compleat_drosophila_methods <- compleat_drosophila %>% group_by(V7) %>% summarise(n_complexes=n())

# yeast
compleat_yeast_methods <- compleat_yeast %>% group_by(V7) %>% summarise(n_complexes=n())

## Predicted methods
# drosophila
compleat_drosophila_predicted <- compleat_drosophila %>% filter(V3=="Predicted") %>% group_by(V7) %>% summarise(n_complexes=n())

# yeast
compleat_yeast_predicted <- compleat_yeast %>% filter(V3=="Predicted") %>% group_by(V7) %>% summarise(n_complexes=n())

```

In yeast and fruit fly the methods of prediction are NetworkBlast and CFinder. Calculate the individual distributions between prediction methods and literature.


```{r}
## Distribution of individual methods
# drosophila

compleat_drosophila_Literature <- compleat_drosophila %>% filter(V3=="Literature") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_drosophila_Literature)[1] <- "complex_with_k_proteins"

compleat_drosophila_NetworkBlast <- compleat_drosophila %>% filter(V7=="NetworkBlast") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_drosophila_NetworkBlast)[1] <- "complex_with_k_proteins"

compleat_drosophila_CFinder <- compleat_drosophila %>% filter(V7=="CFinder") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_drosophila_CFinder)[1] <- "complex_with_k_proteins"

#yeast
compleat_yeast_Literature <- compleat_yeast %>% filter(V3=="Literature") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_yeast_Literature)[1] <- "complex_with_k_proteins"

compleat_yeast_NetworkBlast <- compleat_yeast %>% filter(V7=="NetworkBlast") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_yeast_NetworkBlast)[1] <- "complex_with_k_proteins"

compleat_yeast_CFinder <- compleat_yeast %>% filter(V7=="CFinder") %>% group_by(V2) %>% summarise(n_complex_with_k_proteins=n())
colnames(compleat_yeast_CFinder)[1] <- "complex_with_k_proteins"

```

Summary statistics.
```{r}
summary(compleat_drosophila_NetworkBlast)
summary(compleat_yeast_NetworkBlast)

```

Plot drosophila.

```{r}
ggplot()+
  geom_point(data = compleat_drosophila_Literature, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="Literature"))+
  geom_point(data = compleat_drosophila_CFinder, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="CFinder"))+
  geom_point(data = compleat_drosophila_NetworkBlast, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="NetworkBlast"))+
  ggtitle("Distribution of Proteins in Fruit fly's Protein Complexes")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Literature"="dodgerblue", "CFinder"= "springgreen", "NetworkBlast"= "orchid2"), name="Methods")+
  labs(x="number of proteins", y="number of complexes")+
  theme_bw()
```

Cumulative distribution

```{r}


compleat_drosophila_predicted2 <- compleat_drosophila %>% filter(V7=="NetworkBlast" | V7=="CFinder")

random_dist <- as.data.frame(x=runif(n = 5000,min = 3,max = 189))
colnames(random_dist) <- "Random"

ggplot()+ 
  stat_ecdf(data = compleat_drosophila, aes(x = V2, colour=V3),geom = "step")+
  stat_ecdf(data = compleat_drosophila_predicted2, aes(x = V2, colour=V7),geom = "step")+
  stat_ecdf(data = random_dist, aes(x = Random, colour="Random"),geom = "step")+
  ggtitle("Cumulative distribution of Proteins in Fruit fly's Protein Complexes")+
  labs(x="number of proteins", y="cumulative probability")+
  theme_bw()


```

Is the distribution of literature a power law?? 

```{r,cache=TRUE}
library(poweRlaw)
compleat_drosophila_literature2 <- compleat_drosophila %>% filter(V3=="Literature")

lit_power_law <- displ$new(compleat_drosophila_literature2$V2)

parest <- estimate_pars(lit_power_law)

min_est <- estimate_xmin(lit_power_law)

lit_power_law2 <- lit_power_law$setXmin(min_est)

bs <- bootstrap(lit_power_law, no_of_sims = 100, threads = 1)

is_power_law <- bootstrap_p(lit_power_law, no_of_sims=100, threads=2)

#m_m = displ$new(moby)

#is_power_lawtest <- bootstrap_p(m_m, no_of_sims=100, threads=2)

```

It is not power law based on the Kolmogorov Smirnoff test develoveped by Clauset and implemented in the R CRAN package "poweRlaw". The p-value is 0. 

Plot yeast.

```{r}
ggplot()+
  geom_point(data = compleat_yeast_Literature, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="Literature"))+
  geom_point(data = compleat_yeast_CFinder, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="CFinder"))+
  geom_point(data = compleat_yeast_NetworkBlast, aes(x = complex_with_k_proteins, y = n_complex_with_k_proteins, color="NetworkBlast"))+
  ggtitle("Distribution of Proteins in Yeast's Protein Complexes")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Literature"="dodgerblue", "CFinder"= "springgreen", "NetworkBlast"= "orchid2" ), name="Method")+
  labs(x="number of proteins", y="number of complexes")+
  theme_bw()
```

This is because NetworkBlast has an inherent bias towards small sized complexes or is the result of a faulty interpetation in the article?

# Distribution of the membership of proteins in complexes


```{r}
# drosophila
compleat_drosophila2 <- compleat_drosophila[,c(1,3,7,12)]
drosophila_splitted_complexes <- strsplit(x = as.character(compleat_drosophila2$V12), " ")

compleat_drosophila_long <- data.frame(Complex = rep.int(compleat_drosophila2$V1, sapply(drosophila_splitted_complexes, length)),Method =  rep.int(compleat_drosophila2$V3, sapply(drosophila_splitted_complexes, length)), Tool = rep.int(compleat_drosophila2$V7, sapply(drosophila_splitted_complexes, length)), Protein = unlist(drosophila_splitted_complexes)) # create data frame with 2 columns (Complexes,Proteins). The first column repeat each element of the complexes with the length of the proteins from the list of proteins per complex. Then in the second column put all the proteins of the list (unlist).

# yeast
compleat_yeast2 <- compleat_yeast[,c(1,3,7,12)]
yeast_splitted_complexes <- strsplit(x = as.character(compleat_yeast2$V12), " ")

compleat_yeast_long <- data.frame(Complex = rep.int(compleat_yeast2$V1, sapply(yeast_splitted_complexes, length)), Method =  rep.int(compleat_yeast2$V3, sapply(yeast_splitted_complexes, length)), Tool = rep.int(compleat_yeast2$V7, sapply(yeast_splitted_complexes, length)), Protein = unlist(yeast_splitted_complexes)) 

# human

compleat_homo2 <- compleat_homo[,c(1,5)]
homo_splitted_complexes <- strsplit(x = as.character(compleat_homo2$V5), ";")

compleat_homo_long <- data.frame(Complex = rep.int(compleat_homo2$V1, sapply(homo_splitted_complexes, length)), Protein = unlist(homo_splitted_complexes)) 

```


```{r}
# drosophila
compleat_drosophila_protein_dist <- compleat_drosophila_long %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())
# yeast
compleat_yeast_protein_dist <- compleat_yeast_long %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())
# homo
compleat_homo_protein_dist <- compleat_homo_long %>%  group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

```


```{r}
ggplot()+
  geom_point(data = compleat_drosophila_protein_dist, aes(x = k_complexes, y = n_protein_k_complexes,color="Fruit fly"))+
  geom_point(data = compleat_yeast_protein_dist, aes(x = k_complexes, y = n_protein_k_complexes,color="Yeast")) +
  geom_point(data = compleat_homo_protein_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="Human"))+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Fruit fly"="orange1", "Yeast"= "mediumaquamarine", "Human"="gray48"), name="Organisms")+
  ggtitle("Distribution of Proteins participation in Protein Complexes in Compleat")+
  labs(x="number of complexes of protein", y="number of proteins")+
  theme_bw()
```



```{r}

# drosophila
compleat_drosophila_protein_Literature_dist <- compleat_drosophila_long %>% filter(Method=="Literature") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

compleat_drosophila_protein_NetworkBlast_dist <- compleat_drosophila_long %>% filter(Tool=="NetworkBlast") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

compleat_drosophila_protein_CFinder_dist <- compleat_drosophila_long %>% filter(Tool=="CFinder") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

# yeast
compleat_yeast_protein_Literature_dist <- compleat_yeast_long %>% filter(Method=="Literature") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

compleat_yeast_protein_NetworkBlast_dist <- compleat_yeast_long %>% filter(Tool=="NetworkBlast") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())

compleat_yeast_protein_CFinder_dist <- compleat_yeast_long %>% filter(Tool=="CFinder") %>% group_by(Protein) %>% summarize(k_complexes=n()) %>% group_by(k_complexes) %>% summarise(n_protein_k_complexes=n())



```


```{r}
ggplot()+
  geom_point(data = compleat_drosophila_protein_Literature_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="Literature"))+
  geom_point(data = compleat_drosophila_protein_CFinder_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="CFinder"))+
  geom_point(data = compleat_drosophila_protein_NetworkBlast_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="NetworkBlast"))+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Literature"="dodgerblue", "CFinder"= "springgreen", "NetworkBlast"= "orchid2" ), name="Method")+
  ggtitle("Distribution of Proteins participation in Protein Complexes of Fruit fly")+
  labs(x="number of complexes of protein", y="number of proteins")+
  theme_bw()
```


```{r}
ggplot()+
  geom_point(data = compleat_yeast_protein_Literature_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="Literature"))+
  geom_point(data = compleat_yeast_protein_CFinder_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="CFinder"))+
  geom_point(data = compleat_yeast_protein_NetworkBlast_dist, aes(x = k_complexes, y = n_protein_k_complexes, color="NetworkBlast"))+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="trbl")+
  coord_fixed(ratio = 1)+
  scale_colour_manual(values = c("Literature"="dodgerblue", "CFinder"= "springgreen", "NetworkBlast"= "orchid2" ), name="Method")+
  ggtitle("Distribution of Proteins participation in Protein Complexes of Yeast")+
  labs(x="number of complexes of protein", y="number of proteins")+
  theme_bw()
```

# References







