---
bibliography: [Networks Centralities and Negative Weigths.bib]
mainfont: 12pt
geometry: margin = 1.2in
header-includes:
- \usepackage{placeins}
- \usepackage{fancyhdr}
- \usepackage{setspace}
- \usepackage{chngcntr}
- \usepackage{graphicx}
- \usepackage{microtype}
- \onehalfspacing
- \counterwithin{figure}{section}
- \counterwithin{table}{section}
- \renewcommand{\familydefault}{\sfdefault}
- \usepackage{helvet}
- \usepackage{mathtools}
- \usepackage{amsmath}
- \usepackage{sansmathfonts}
- \newtheorem{theorem}{Theorem}
- \newtheorem{mydef}{Definition}
- \usepackage[utf8]{inputenc}
- \usepackage[greek,english]{babel}
- \usepackage{multirow}
- \usepackage{subcaption}
- \usepackage{wrapfig}
- \usepackage{array}
linkcolor: black
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
  html_document: default
  word_document: default
link-citations: true
urlcolor: blue
citecolor: black
---

---
nocite: 
...
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path = 'Thesis_Essentiality_Drosophila_Signed_Network/Figures/',
                      echo = FALSE, warning = FALSE, message = FALSE)
```

\pagenumbering{gobble}

```{r child = 'titlepage.Rmd'}
```


\newpage
\pagestyle{fancy}

\fancyhead[LE,RO]{}
\fancyhead[LO,RE]{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\pagenumbering{roman}

\fancyhead[CO,CE]{Copyright}
\section*{Copyright}
\addcontentsline{toc}{section}{Copyright}

```{r child = 'copyright.Rmd',eval=T}
```

\newpage
\fancyhead[CO,CE]{Abstract}
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

```{r child = 'abstract.Rmd'}
```

\newpage

\fancyhead[CO,CE]{Table of Contents}
\setcounter{tocdepth}{2}
\tableofcontents

\newpage
\pagenumbering{arabic}

\newpage
\fancyhead[CO,CE]{Introduction}

```{r child = 'chapter1.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Methods}

```{r child = 'chapter2.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Results}

```{r child = 'chapter3.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Discussion}

```{r child = 'chapter4.Rmd'}
```

\FloatBarrier

\newpage
\fancyhead[CO,CE]{List of Figures}
\addcontentsline{toc}{section}{List of Figures}
\listoffigures

\newpage
\fancyhead[CO,CE]{List of Tables}
\addcontentsline{toc}{section}{List of Tables}
\listoftables

\newpage

\fancyhead[CO,CE]{Appendices}
\addcontentsline{toc}{section}{Appendices}

```{r child = 'appendix.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{References}

# References
